APPM report - Draft

### 2022-12-21 paper_version/main.tex
Edited Sections 1-2

We elaborate a numerical method for a three-dimensional hydrodynamic multi-species plasma
model which boils down to an extended Euler-Maxwell system. Our method is inspired by and
extends the one-dimensional scheme from $[$P.~Degond, F.~Deluzet, and D.~Savelief,
Numerical approximation of the Euler-Maxwell model in the quasineutral limit, Journal of
Computational Physics, 231 (4), pp.~1917--1946, 2012$]$, it can cope with large variations
of the so-called Debye length $\lambda_D$ and is robust in the quasi-neutral
singular-perturbation limit $\lambda_D\to 0$, because it enjoys an
\emph{asymptotic-preserving} (AP) property in the sense that in the sense that the limit
$\lambda_D\to 0$ of the scheme yields a viable discretization for the continuous limit
model. The key ingredients of our approach are (i) a discretization of Maxwell's equations
based on primal and dual meshes in the spirit of \emph{discrete exterior calculus} (DEC)
also known the finite integration technique (FIT), (ii) a finite volume method (FVM) for
the fluid equations on the dual mesh, (iii) \emph{mixed implicit-explicit timestepping},
(iv) special no-flux and contact boundary conditions at an outer cut-off boundary, and (v)
additional \emph{stabilization} in the non-conducting region outside the plasma domain
based on direct enforcement of Gauss' law. Numerical tests provide strong evidence
confirming the AP property of the new method.

@@

% \mycomment{Roman: more info from a physical point of view?} R.H. 19.12. Extended paragraph
The starting point for this work was the desire to numerically simulate the formation and evolution of electric arcs at atmospheric pressures. We rely on a widely accepted description by means of a hydrodynamic multi-species plasma model, which boils down to the Maxwell-Euler system. The arc phenomenon covers a wide range of plasma regimes and the numerical model should be able to deal with all of them seamlessly and simultaneously. Thus in this work, as in \cite{degond_2012}, the focus is on \emph{asymptotic-preserving} (AP) discretization in the quasi-neutral limit $\lambda \rightarrow 0$, where $\lambda$ is the rescaled Debye length, see Section \ref{sec:rescaling} for the underlying scaling arguments and \cite{degond_2012,degond_2017} for a discussion of the importance of the AP property of numerical plasma models.

@@

The quasi-neutral limit of the Maxwell-Euler system leads to a singularly perturbed problem, that is, the limiting PDE system changes its type\footnote{The Maxwell-Euler system transits from hyperbolic to a mixed type}. This poses a challenge for simulations in settings encompassing different regimes. Singularly perturbed problems occur in many physical models, e.g. vanishing viscosity in the Navier-Stokes equations \cite{Kato_1984}, zero-relaxation-limit of kinetic equations \cite{jin_2010} and zero-Mach-limit of compressible flows \cite{degond_2007, haack_2010}. It is desirable that numerical schemes are valid for all ranges of scaled parameters. These schemes are then said to be \emph{asymptotic-preserving} (AP). We discretize a parametric problem $P^\lambda$, which converges to its limit $P^0$ as the perturbed parameter $\lambda \rightarrow 0$, by the scheme $P^\lambda_h$ where $h$ denotes some discretization parameter, e.g. the mesh size, see also the illustration by a commuting diagram in \cite[][Fig. 1]{jin_2010}. The AP property requires the uniform convergence of $P^\lambda_h$ to the $P^\lambda$ as $\lambda\rightarrow0$. Readers are referred to the comprehensive reviews in \cite{jin_2010, degond_2017} for more details about AP schemes.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
    \coordinate (O1) at (0,0);
    \draw[violet] (O1) circle (1) node[blue] {Electrode};
    \fill[blue, opacity=0.2] (O1) circle (1);
    \draw[very thick] (O1) circle (1.5);
    \coordinate (O2) at (4.5,1);
    \draw[violet] (O2) circle (1);
    \node at (O2) {\textcolor{red}{Electrode}};
    \node at (barycentric cs:O1=1,O2=1) {Plasma}; 
    \fill[red, opacity=0.2] (O2) circle (1);
    \coordinate (S1) at ([shift={(-0.196,0.98)}]O1);
    \coordinate (S2) at ([shift={(-0.196,0.98)}]O2);
    \coordinate (P1) at ([shift={(0.196,-0.98)}]O1);
    \coordinate (P2) at ([shift={(0.196,-0.98)}]O2);
    \draw[violet] (S1) -- (S2);
    \draw[violet] (P1) -- (P2);
    \coordinate (SS1) at ([shift={({-0.196*1.5},{0.98*1.5})}]O1);
    \coordinate (SS2) at ([shift={({-0.196*1.5},{0.98*1.5})}]O2);
    \coordinate (PP1) at ([shift={({0.196*1.5},{-0.98*1.5})}]O1);
    \coordinate (PP2) at ([shift={({0.196*1.5},{-0.98*1.5})}]O2);
    \draw[very thick] (SS1) -- (SS2);
    \draw[very thick] (PP1) -- (PP2); 
    \draw[black, very thick] (SS2) arc (101.31:180+101.31:1.5);
    \draw[very thick] (PP2) arc (101.31-180:101.31:1.5);
    \draw[-to] (barycentric cs:SS1=1,SS2=1,S1=1,S2=1) to[out=90,in=0] ([shift={(-0.5,0.5)}]barycentric cs:SS1=1,SS2=1,S1=1,S2=1) node[left, gray] {Insulator};
    \begin{scope}[shift={(5.5,-1.4)}]
    \draw[blue, -to] (0,0) -- (0.65,0.12) node[right]{\small $z$};
    \draw[blue, -to] (0,0) -- (0,0.7) node[above]{\small $x$};
    \draw[blue, -to] (0,0) -- (0.7,0) node[below]{\small $y$};
    \end{scope}
    \end{tikzpicture}
\caption{The domain geometry}
\label{fig:domain_geometry}
\end{figure}

@@

\subsection{The Maxwell-Euler System}
Plasmas consist of electrically charged and (possibly) neutral particle species that can be treated as fluids. That is, they may be described by the Navier-Stokes equations augmented with source terms for species generation (ionization, recombination), forces (Lorentz force and interspecies friction), and energy (thermal relaxation among species, radiative heat transfer, Ohmic heating). For more details on fluid models of plasmas, see \cite{chen2016, remi_2014}.

The electromagnetic effects in plasmas are governed by Maxwell's equations:
\begin{subequations}
\begin{align}
    \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0, \label{equ:maxwell_faraday} \\ 
    \partial_t \mathbf{D} - \nabla \times \mathbf{H} &= -\mathbf{J}, \label{equ:maxwell_ampere} \\
    \nabla \cdot \mathbf{B} &= 0,  \label{equ:maxwell_gauss_B}\\
    \nabla \cdot \mathbf{D} &= \rho, \label{equ:maxwell_gauss_D}
\end{align}
\end{subequations}
where the involved quantities\footnote{SI units in square brackets.} are the magnetic flux density $\mathbf{B}[\frac{\text{Wb}}{\text{m}^2}]$, the electric field $\mathbf{E}[\frac{\text{V}}{\text{m}}]$, the electric displacement field $\mathbf{D}[\frac{\text{C}}{\text{m}^2}]$, the magnetic field $\mathbf{H}[\frac{\text{A}}{\text{m}}]$, the electric current density $\mathbf{J}[\frac{\text{A}}{\text{m}^2}]$ and the electric charge density $\rho[\frac{\text{C}}{\text{m}^3}]$. These equations are known as Faraday's law (\ref{equ:maxwell_faraday}), Amp\`{e}re's law (\ref{equ:maxwell_ampere}), Gauss's law for magnetism (\ref{equ:maxwell_gauss_B}) and Gauss's law for electric field (\ref{equ:maxwell_gauss_D}). For simplicity, we only consider isotropic and homogeneous media. Namely, we rely on linear constitutive relations (i.e. material laws):
\begin{equation} \label{equ:material_law}
    \mathbf{D} = \epsilon \mathbf{E}, \ \ \ \mathbf{B} = \mu\mathbf{H}, 
\end{equation}
where $\epsilon[\frac{\text{F}}{\text{m}}]$ and $\mu[\frac{\text{N}}{\text{A}^{2}}]$ are the permittivity and permeability and assumed to be constant. By taking the time derivative of (\ref{equ:maxwell_gauss_D}) and the divergence of (\ref{equ:maxwell_ampere}), it is easy to see that (\ref{equ:maxwell_ampere}) and (\ref{equ:maxwell_gauss_D}) implies the charge conservation
\begin{equation} \label{equ:charge_continuity}
    \partial_t\rho + \nabla \cdot \mathbf{J} = 0.
\end{equation}
Besides, by taking the divergence of equations (\ref{equ:maxwell_faraday}) and (\ref{equ:maxwell_ampere}) together with the relation (\ref{equ:charge_continuity}), it is possible to verify that Gauss's laws (\ref{equ:maxwell_gauss_B}) (\ref{equ:maxwell_gauss_D}) are consequences of (\ref{equ:maxwell_faraday}) and (\ref{equ:maxwell_ampere}) provided that (\ref{equ:maxwell_gauss_B}) (\ref{equ:maxwell_gauss_D}) are satisfied at initial time. In other words, Gauss's laws (\ref{equ:maxwell_gauss_B}) (\ref{equ:maxwell_gauss_D}) can be thought of as initial conditions.

Each species in a plasma can be modelled as a compressible fluid through a set of Euler equations augmented by electromagnetic forces as well as other processes, e.g. collisions and recombination:
\begin{align} \label{equ:euler_}
    \partial_t
    \begin{pmatrix}
    n_* \\
    m_*n_* \mathbf{u}_* \\
    m_*n_* e_*
    \end{pmatrix}
    + \nabla \cdot
    \begin{pmatrix}
    n_* \mathbf{u}_* \\
    m_*n_* \mathbf{u}_* \otimes \mathbf{u}_* + p_*\mathbb{I} \\
    m_*n_* h_* \mathbf{u}_*
    \end{pmatrix}
    =
    \begin{pmatrix}
    0 \\
    q_*n_*(\mathbf{E} + \mathbf{u}_* \times \mathbf{B}) \\
    \mathbf{J}_* \cdot \mathbf{E}
    \end{pmatrix}
    +
    \begin{pmatrix}
    \Gamma_* \\
    \mathbf{R}_* \\
    Q_*
    \end{pmatrix}.
\end{align}
The subscript $*\in \{e, i, \cdots\}$ labels electrons, ions or other species. The involved fluid quantities are: number density $n_*[\frac{1}{\text{m}^3}]$, particle mass $m_*[\text{kg}]$,  electric charge $q_*[\text{C}]$, velocity $\mathbf{u}_*[\frac{\text{m}}{\text{s}}]$, pressure $p_*[\frac{\text{N}}{\text{m}^2}]$, specific total energy $e_* := \frac{1}{2}|\mathbf{u}_*|^2 + \frac{p_*}{m_*n_*(\gamma - 1)}[\frac{\text{J}}{\text{kg}}]$, specific total enthalpy $h_* := e_* + \frac{p_*}{m_*n_*}[\frac{\text{J}}{\text{kg}}]$ and current density $\mathbf{J}_* := q_*n_*\mathbf{u}_*[\frac{\text{A}}{\text{m}^2}]$. All the species are assumed to be ideal gases with heat capacity ratio $\gamma=5/3$ for monatomic gases. The left-hand side of the equations describes the convection of plasma fluids. The first term at the right-hand side takes into account the Lorentz force in the momentum equation and Joule heating in the energy equation. $\Gamma_*, \mathbf{R}_*, Q_*$ are generic terms due to effects of inter-species collision, ionization, recombination, etc.

For the sake of simplicity, we only consider ion-electron collision in this work, but extension to more species is straightforward. Such collisions can be well modelled by friction terms (\citep[][Sec. 5.6.2]{chen2016})
\begin{equation} \label{equ:collision}
    \mathbf{R}_{e}^{\text{coll}} = \frac{\pi Ze^4m_e^{1/2}}{(4\pi\epsilon)^2(k_BT_e)^{3/2}}\text{ln}\Lambda n_en_i(\mathbf{u}_i - \mathbf{u}_e) = - \mathbf{R}_{i}^{\text{coll}},  
\end{equation}
which is the collision force acting on electrons. ln$\Lambda$ is called the \emph{Coulomb logarithm} and is usually taken as a constant \cite{chen2016}. Because of the conservation of momentum, the collision term for ions $\mathbf{R}_{i}^{\text{coll}}$ is nothing but the negative of $\mathbf{R}_{e}^{\text{coll}}$. In this work, the other effects, e.g. ionization and recombination, are not considered. For full details about the modeling of these physical processes, readers are referred to \cite[][Sec. 3.2]{fuchs_2021}.

Combination of the two sets of equations (\ref{equ:maxwell_faraday}) to (\ref{equ:maxwell_gauss_D}) and (\ref{equ:euler_}) leads to the governing equations for a multi-species plasma model. The coupling of the two subsystems is due to the fact that the flow of charged particles induces electric currents:
\begin{equation} \label{equ:maxwell_euler_coupling}
    \rho = \sum_* q_*n_*, \ \ \  \mathbf{J} = \sum_* q_*n_*\mathbf{u}_*. 
\end{equation}

\subsection{Rescaling} \label{sec:rescaling}
In rescaling, variables in equations are replaced by the ratio of original quantities and constant reference quantities, which transforms the system into dimensionless form. In this part, the rescaling procedure mostly follows that of \cite[][Sec. 2.2]{degond_2017}.

First, a reference value for each quantity is prescribed. We define $\overline{x}$ and $\overline{t}$ as space and time scale of interest, by which the reference velocity $\overline{v}$ is defined as $\overline{x}/\overline{t}$. The typical magnitudes of electromagnetic fields are $\overline{E}$ and $\overline{B}$. The reference magnitudes of fluid quantities are denoted by $\overline{u}$ for plasma drift velocity, $\overline{n}$ for plasma number density, $\overline{T}$ for plasma temperature. On top of that, the reference pressure is set to $\overline{n}k_B\overline{T}$ with $k_B$ being the Boltzmann constant; the reference value for the specific total energy and enthalpy is set to $\overline{u}^2$. Also taking into account the physical constants, i.e. the reference permittivity $\overline{\epsilon}$, the reference permeability $\overline{\mu}$, the light speed $c^2 = (\overline{\epsilon}\overline{\mu})^{-1}$, Boltzmann constant $k_B$, the elementary charge $e$, and ion mass $m_i$, we introduce the following dimensionless parameters:
\begin{align*} 
    \xi &:= \frac{\overline{v}}{c} \text{ the velocity of interest \emph{w.r.t.}\footnotemark[1] the light speed}, &\\
    \zeta &:= \frac{\overline{u}}{\overline{v}} \text{ the plasma velocity \emph{w.r.t.} the speed of interest}, &\\
    M &:= \frac{\overline{u}}{v_i^{\text{th}}} \text{ the plasma drift velocity \emph{w.r.t.} the thermal speed of ions, where } v_i^{\text{th}} := \left(\frac{k_B \overline{T}}{m_i}\right)^{1/2}, &\\
    \eta &:= \frac{e\overline{E}\overline{x}}{m_i\overline{u}^2} \text{ the electric energy \emph{w.r.t.} the kinetic energy of ions}, &\\
    \beta &:= \frac{\overline{v}\overline{B}}{\overline{E}} \text{ the induced electric field \emph{w.r.t.} the total electric field}, &\\
    \lambda &:= \frac{\lambda_D}{\overline{x}} \text{ the dimensionless Debye length, where }\lambda_D := \left(\frac{\epsilon k_B\overline{T}}{e^2\overline{n}}\right)^{1/2}, &\\
    \varepsilon_*^2 &:= \frac{m_*}{m_i} \text{ the species mass \emph{w.r.t.} the ion mass},\ \  * \in \{i,e,\cdots\}, &\\
    g &:= \frac{1}{\overline{n}\lambda_D^3} \text{ the plasma parameter\footnotemark[2]}.&
\end{align*}
\footnotetext[1]{Abbreviation of \emph{with respect to}.}
\footnotetext[2]{It is basically the reciprocal of the number of particles per Debye sphere, see \cite[][Sec. 1.1]{gibbon2020}. Classical plasma theory assumes $g \ll 1$.}

@@

Special treatment is needed for the collision term. One could argue that the occurrence of $\lambda^{-1}$ in (\ref{equ:rescaling_friction}) could lead to singular perturbation of the Euler system as $\lambda \rightarrow 0$. However, the validity of such a collision model is doubtful when $\lambda \rightarrow 0$. From the physical viewpoint, the shielding effect characterized by the rescaled Debye length $\lambda$ makes the impact parameter no larger than $\lambda$. The $\lambda \rightarrow 0$ limit implies that the collisions happen in the atomic length scale, in which quantum-mechanical effects must be taken into effect \cite[][pp. 144-148]{frank_1972}. Therefore, we eliminate the perturbation parameter in the friction term by assuming $g \sim \lambda$ and keep the prefactor $\alpha^{\text{coll}}$ of the collision term constant throughout the simulation.

\begin{remark}
    The case when there is an asymptotic parameter at the denominator of the friction term is of interest, too. Such problem is called relaxation limit, see \cite{peng_2011, wasiolek_2016, Li_2021, Chen_2000}.
\end{remark}

In summary, we make the following choices \cite[][sec 2.2]{degond_2017}
\begin{align*}
    \xi \sim g \sim \lambda \in [0,1], \ \ \ \zeta \sim M \sim \eta \sim \beta \sim \mathcal{O}(1).
\end{align*}

>>>>

We elaborate a numerical method for a three-dimensional hydrodynamic multi-species plasma
model which boils down to an extended Euler-Maxwell system. Our method is inspired by and
extends the one-dimensional scheme from $[$P.~Degond, F.~Deluzet, and D.~Savelief,
\emph{Numerical approximation of the Euler-Maxwell model in the quasineutral limit}, Journal of
Computational Physics, 231 (4), pp.~1917--1946, 2012$]$. It can cope with large variations
of the so-called Debye length $\lambda_D$ and is robust in the quasi-neutral
singular-perturbation limit $\lambda_D\to 0$, because it enjoys an
\emph{asymptotic-preserving} (AP) property in the sense that in the sense that the limit
$\lambda_D\to 0$ of the scheme yields a viable discretization for the continuous limit
model. The key ingredients of our approach are (i) a discretization of Maxwell's equations
based on primal and dual meshes in the spirit of \emph{discrete exterior calculus} (DEC)
also known the finite integration technique (FIT), (ii) a finite volume method (FVM) for
the fluid equations on the dual mesh, (iii) \emph{mixed implicit-explicit timestepping},
(iv) special no-flux and contact boundary conditions at an outer cut-off boundary, and (v)
additional \emph{stabilization} in the non-conducting region outside the plasma domain
based on direct enforcement of Gauss' law. Numerical tests provide strong evidence
confirming the AP property of the new method.

@@

% \mycomment{Roman: more info from a physical point of view?} R.H. 19.12. Extended paragraph
The starting point for this work was the desire to numerically simulate the formation and
evolution of electric arcs at atmospheric pressures. Following common practise, we rely on
a mathematical description by means of a hydrodynamic multi-species plasma model, which
boils down to an extended Maxwell-Euler system. The arc phenomenon covers a wide range of
plasma regimes and, thus, the design goal was a numerical model capable of dealing
with all of them seamlessly and simultaneously. Thus in this work, as in
\cite{degond_2012}, the focus is on \emph{asymptotic-preserving} (AP) discretization in
the quasi-neutral limit $\lambda \rightarrow 0$, where $\lambda$ is the rescaled Debye
length, see Section \ref{sec:rescaling} for the underlying scaling arguments and
\cite{degond_2012,degond_2017} for a discussion of the importance of the AP property of
numerical plasma models.

@@

The quasi-neutral limit of the Maxwell-Euler system leads to a singularly perturbed
problem, that is, the limiting PDE system changes its type\footnote{The type of the
  Maxwell-Euler system switches from hyperbolic to mixed for $\lambda_{D}\to 0$}. This
poses a challenge for simulations in settings encompassing different regimes. We remind
that singularly perturbed problems occur in many physical models, e.g., in the case of
vanishing viscosity in the Navier-Stokes equations \cite{Kato_1984}, the
zero-relaxation-limit of kinetic equations \cite{jin_2010}, and the zero-Mach-number limit
of compressible flows \cite{degond_2007, haack_2010}. Throughout, it is essential that
numerical schemes remain valid even if crucial model parameters approach the limit. These
schemes are then said to be \emph{asymptotic-preserving} (AP): Let us assume that we
discretize a parameter($\lambda$)-dependent model $P^\lambda$, which converges to a limit
$P^0$ as $\lambda \rightarrow 0$, by the scheme $P^\lambda_h$ where $h$ denotes some
discretization parameter, e.g., the mesh size. The AP property amounts to uniform
convergence of $P^\lambda_h$ to the $P^\lambda$ as $\lambda\rightarrow0$. Readers are
referred to the illustration by a commuting diagram in \cite[][Fig. 1]{jin_2010} and, more
generally, to the comprehensive reviews \cite{jin_2010, degond_2017} for more details
about AP schemes.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=1.2]
    \coordinate (O1) at (0,0);
    \draw[violet] (O1) circle (1) node[blue] {Electrode};
    \fill[blue, opacity=0.2] (O1) circle (1);
    \draw[very thick] (O1) circle (1.5);
    \coordinate (O2) at (4.5,1);
    \draw[violet] (O2) circle (1);
    \node at (O2) {\textcolor{red}{Electrode}};
    \node at (barycentric cs:O1=1,O2=1) {Plasma}; 
    \fill[red, opacity=0.2] (O2) circle (1);
    \coordinate (S1) at ([shift={(-0.196,0.98)}]O1);
    \coordinate (S2) at ([shift={(-0.196,0.98)}]O2);
    \coordinate (P1) at ([shift={(0.196,-0.98)}]O1);
    \coordinate (P2) at ([shift={(0.196,-0.98)}]O2);
    \draw[violet] (S1) -- (S2);
    \draw[violet] (P1) -- (P2);
    \coordinate (SS1) at ([shift={({-0.196*1.5},{0.98*1.5})}]O1);
    \coordinate (SS2) at ([shift={({-0.196*1.5},{0.98*1.5})}]O2);
    \coordinate (PP1) at ([shift={({0.196*1.5},{-0.98*1.5})}]O1);
    \coordinate (PP2) at ([shift={({0.196*1.5},{-0.98*1.5})}]O2);
    \draw[very thick] (SS1) -- (SS2);
    \draw[very thick] (PP1) -- (PP2); 
    \draw[black, very thick] (SS2) arc (101.31:180+101.31:1.5);
    \draw[very thick] (PP2) arc (101.31-180:101.31:1.5);
    \draw[-to] (barycentric cs:SS1=1,SS2=1,S1=1,S2=1) to[out=90,in=0] ([shift={(-0.5,0.5)}]barycentric cs:SS1=1,SS2=1,S1=1,S2=1) node[left, gray] {Insulator};
    \begin{scope}[shift={(5.5,-1.4)}]
    \draw[blue, -to] (0,0) -- (0.65,0.12) node[right]{\small $z$};
    \draw[blue, -to] (0,0) -- (0,0.7) node[above]{\small $x$};
    \draw[blue, -to] (0,0) -- (0.7,0) node[below]{\small $y$};
    \end{scope}
    \end{tikzpicture}
    \caption{The special geometric setting suggested by the simulation of electric arcs,
      for which we discuss the new method.}
\label{fig:domain_geometry}
\end{figure}

@@

\subsection{The Maxwell-Euler System}

Plasmas consist of electrically charged and (possibly) neutral particle species and we
suppose that they can be treated as continua (fluids). That is, they may be described by the
Navier-Stokes equations augmented with source terms for species generation (ionization,
recombination), forces (Lorentz force and interspecies friction), and energy (thermal
relaxation among species, radiative heat transfer, Ohmic heating). For more details on
fluid models of plasmas, see \cite{chen2016, remi_2014}.

The electromagnetic effects in plasmas are governed by Maxwell's equations:
\begin{subequations}
\begin{align}
    \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0, \label{equ:maxwell_faraday} \\ 
    \partial_t \mathbf{D} - \nabla \times \mathbf{H} &= -\mathbf{J}, \label{equ:maxwell_ampere} \\
    \nabla \cdot \mathbf{B} &= 0,  \label{equ:maxwell_gauss_B}\\
    \nabla \cdot \mathbf{D} &= \rho, \label{equ:maxwell_gauss_D}
\end{align}
\end{subequations}
where the involved quantities\footnote{SI units in square brackets.} are the magnetic flux
density $\mathbf{B}[\frac{\text{Wb}}{\text{m}^2}]$, the electric field
$\mathbf{E}[\frac{\text{V}}{\text{m}}]$, the electric displacement field
$\mathbf{D}[\frac{\text{C}}{\text{m}^2}]$, the magnetic field
$\mathbf{H}[\frac{\text{A}}{\text{m}}]$, the electric current density
$\mathbf{J}[\frac{\text{A}}{\text{m}^2}]$ and the electric charge density
$\rho[\frac{\text{C}}{\text{m}^3}]$. These equations are known as Faraday's law
(\ref{equ:maxwell_faraday}), Amp\`{e}re's law (\ref{equ:maxwell_ampere}), Gauss's law for
magnetism (\ref{equ:maxwell_gauss_B}) and Gauss's law for electric field
(\ref{equ:maxwell_gauss_D}). For simplicity, we only consider isotropic and homogeneous
media. Namely, we rely on linear constitutive relations (material laws):
\begin{equation} \label{equ:material_law}
    \mathbf{D} = \epsilon \mathbf{E}\quad,\quad\mathbf{B} = \mu\mathbf{H}, 
\end{equation}
where $\epsilon[\frac{\text{F}}{\text{m}}]$ and $\mu[\frac{\text{N}}{\text{A}^{2}}]$ are
the permittivity and permeability and assumed to be constant. By taking the time
derivative of (\ref{equ:maxwell_gauss_D}) and the divergence of
(\ref{equ:maxwell_ampere}), it is easy to see that (\ref{equ:maxwell_ampere}) and
(\ref{equ:maxwell_gauss_D}) implies the charge conservation
\begin{equation} \label{equ:charge_continuity}
    \partial_t\rho + \nabla \cdot \mathbf{J} = 0.
\end{equation}
Besides, by taking the divergence of equations (\ref{equ:maxwell_faraday}) and
(\ref{equ:maxwell_ampere}) together with the relation (\ref{equ:charge_continuity}), it is
possible to verify that Gauss's laws (\ref{equ:maxwell_gauss_B})
(\ref{equ:maxwell_gauss_D}) are consequences of (\ref{equ:maxwell_faraday}) and
(\ref{equ:maxwell_ampere}) provided that (\ref{equ:maxwell_gauss_B})
(\ref{equ:maxwell_gauss_D}) are satisfied at initial time. In other words, Gauss's laws
(\ref{equ:maxwell_gauss_B}) (\ref{equ:maxwell_gauss_D}) can be thought of as initial
conditions.

We model each species in the plasma as a compressible fluid through a set of \emph{Euler
  equations} augmented by electromagnetic forces as well as other processes,
e.g. collisions and recombination:
\begin{align} \label{equ:euler_}
    \partial_t
    \begin{pmatrix}
    n_* \\
    m_*n_* \mathbf{u}_* \\
    m_*n_* e_*
    \end{pmatrix}
    + \nabla \cdot
    \begin{pmatrix}
    n_* \mathbf{u}_* \\
    m_*n_* \mathbf{u}_* \otimes \mathbf{u}_* + p_*\mathbb{I} \\
    m_*n_* h_* \mathbf{u}_*
    \end{pmatrix}
    =
    \begin{pmatrix}
    0 \\
    q_*n_*(\mathbf{E} + \mathbf{u}_* \times \mathbf{B}) \\
    \mathbf{J}_* \cdot \mathbf{E}
    \end{pmatrix}
    +
    \begin{pmatrix}
    \Gamma_* \\
    \mathbf{R}_* \\
    Q_*
    \end{pmatrix}.
\end{align}
The subscript $*\in \{e, i, \cdots\}$ labels electrons, ions or other species. The
involved fluid quantities are: number density $n_*[\frac{1}{\text{m}^3}]$, particle mass
$m_*[\text{kg}]$, electric charge $q_*[\text{C}]$, velocity
$\mathbf{u}_*[\frac{\text{m}}{\text{s}}]$, pressure $p_*[\frac{\text{N}}{\text{m}^2}]$,
specific total energy
$e_* := \frac{1}{2}|\mathbf{u}_*|^2 + \frac{p_*}{m_*n_*(\gamma -
  1)}[\frac{\text{J}}{\text{kg}}]$, specific total enthalpy
$h_* := e_* + \frac{p_*}{m_*n_*}[\frac{\text{J}}{\text{kg}}]$ and current density
$\mathbf{J}_* := q_*n_*\mathbf{u}_*[\frac{\text{A}}{\text{m}^2}]$. All species are
assumed to be ideal gases with heat capacity ratio $\gamma=5/3$ for monatomic gases. The
left-hand side of the equations describes the convection of plasma fluids. The first term
at the right-hand side takes into account the Lorentz force in the momentum equation and
Joule heating in the energy equation. $\Gamma_*, \mathbf{R}_*, Q_*$ are generic terms due
to effects of inter-species collision, ionization, recombination, etc.

Combining the two sets of equations (\ref{equ:maxwell_faraday}) to
(\ref{equ:maxwell_gauss_D}) and (\ref{equ:euler_}) leads to the governing equations for a
multi-species plasma model. The coupling of the two subsystems is due to the fact that the
flow of charged particles induces electric currents:
\begin{equation} \label{equ:maxwell_euler_coupling}
    \rho = \sum_* q_*n_*\;, \quad \mathbf{J} = \sum_* q_*n_*\mathbf{u}_*. 
\end{equation}

\begin{remark}
  \label{rem:coll}
  For the sake of simplicity, we only consider ion-electron collision in this work, but
extension to more species is straightforward. Such collisions can be well modelled by
friction terms, for instance \citep[][Sec. 5.6.2]{chen2016}
\begin{equation} \label{equ:collision}
    \mathbf{R}_{e}^{\text{coll}} = \frac{\pi Ze^4m_e^{1/2}}{(4\pi\epsilon)^2(k_BT_e)^{3/2}}\text{ln}\Lambda n_en_i(\mathbf{u}_i - \mathbf{u}_e) = - \mathbf{R}_{i}^{\text{coll}},  
\end{equation}
which is the collision force acting on electrons. ln$\Lambda$ is called the \emph{Coulomb
  logarithm} and is usually taken as a constant \cite{chen2016}. Because of the
conservation of momentum, the collision term for ions $\mathbf{R}_{i}^{\text{coll}}$ is
nothing but the negative of $\mathbf{R}_{e}^{\text{coll}}$. In this work, the other
effects, e.g. ionization and recombination, are not considered. For full details about the
modeling of these physical processes, readers are referred to
\cite[][Sec. 3.2]{fuchs_2021}.
\end{remark}

\subsection{Rescaling}
\label{sec:rescaling}
% In rescaling, variables in equations are replaced by the ratio of original quantities and constant reference quantities, which transforms the system into dimensionless form. In this part, the rescaling procedure mostly follows that of \cite[][Sec. 2.2]{degond_2017}.

We arrive at non-dimensional equations by following the rescaling procedure of
\cite[][Sec. 2.2]{degond_2017}.  First, a reference value for each quantity is
prescribed. We define $\overline{x}$ and $\overline{t}$ as space and time scale of
interest, by which the reference velocity $\overline{v}$ is defined as
$\overline{x}/\overline{t}$. The typical magnitudes of electromagnetic fields are
$\overline{E}$ and $\overline{B}$. The reference magnitudes of fluid quantities are
denoted by $\overline{u}$ for plasma drift velocity, $\overline{n}$ for plasma number
density, $\overline{T}$ for plasma temperature. On top of that, the reference pressure is
set to $\overline{n}k_B\overline{T}$ with $k_B$ being the Boltzmann constant; the
reference value for the specific total energy and enthalpy is set to
$\overline{u}^2$. Also taking into account the physical constants, i.e. the reference
permittivity $\overline{\epsilon}$, the reference permeability $\overline{\mu}$, the light
speed $c^2 = (\overline{\epsilon}\overline{\mu})^{-1}$, Boltzmann constant $k_B$, the
elementary charge $e$, and ion mass $m_i$, we introduce the following dimensionless
parameters:
\begin{align*} 
  \xi &:= \frac{\overline{v}}{c} \text{ the velocity of interest \emph{w.r.t.}\footnotemark[1] the light speed}, &\\
  \zeta &:= \frac{\overline{u}}{\overline{v}} \text{ the plasma velocity \emph{w.r.t.} the speed of interest}, &\\
  M &:= \frac{\overline{u}}{v_i^{\text{th}}} \text{ the plasma drift velocity \emph{w.r.t.} the thermal speed of ions, where } v_i^{\text{th}} := \left(\frac{k_B \overline{T}}{m_i}\right)^{1/2}, &\\
  \eta &:= \frac{e\overline{E}\overline{x}}{m_i\overline{u}^2} \text{ the electric energy \emph{w.r.t.} the kinetic energy of ions}, &\\
  \beta &:= \frac{\overline{v}\overline{B}}{\overline{E}} \text{ the induced electric field \emph{w.r.t.} the total electric field}, &\\
  \lambda &:= \frac{\lambda_D}{\overline{x}} \text{ the dimensionless Debye length, where }\lambda_D := \left(\frac{\epsilon k_B\overline{T}}{e^2\overline{n}}\right)^{1/2}, &\\
  \varepsilon_*^2 &:= \frac{m_*}{m_i} \text{ the species mass \emph{w.r.t.} the ion mass},\ \  * \in \{i,e,\cdots\}, &\\
  g &:= \frac{1}{\overline{n}\lambda_D^3} \text{ the plasma parameter\footnotemark[2]}.&
\end{align*}
\footnotetext[1]{Abbreviation of \emph{with respect to}.}
\footnotetext[2]{It is basically the reciprocal of the number of particles per Debye sphere, see \cite[][Sec. 1.1]{gibbon2020}. Classical plasma theory assumes $g \ll 1$.}

@@

\begin{remark}
Special treatment is needed for the collision term. One could argue that the occurrence of
$\lambda^{-1}$ in (\ref{equ:rescaling_friction}) could lead to singular perturbation of
the Euler system as $\lambda \rightarrow 0$. However, the validity of such a collision
model is doubtful when $\lambda \rightarrow 0$. From the physical viewpoint, the shielding
effect characterized by the rescaled Debye length $\lambda$ limits the impact parameters
to values $\sim\lambda$. The $\lambda \rightarrow 0$ limit implies that the collisions
happen on the atomic length scale, in which quantum-mechanical effects must be taken into
account \cite[][pp. 144-148]{frank_1972}. Therefore, we eliminate the perturbation
parameter in the friction term by assuming $g \sim \lambda$ and keep the prefactor
$\alpha^{\text{coll}}$ of the collision term constant throughout the simulation.

The case when there is an asymptotic parameter in the denominator of the friction term is
of interest, too. This situation is known as relaxation limit, see \cite{peng_2011,
  wasiolek_2016, Li_2021, Chen_2000}.
\end{remark}
